Rapid shortening at the eastern margin of the Tibetan plateau prior to the 2008 $M_\textrm{w}$ 7.9 Wenchuan earthquake

### 2014-10-27 paper.tex
Cite Cooke.

To model both the tectonic and earthquake cycle processes contributing to these interseismic GPS data, we use a quasistatic elastic boundary element software under active development. An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems. Previous boundary element methods in the earth sciences \citep{Thomas1993, Maerten2014a} have been based on analytic integrations of rectangular or triangular slip surfaces \citep{Okada1992, Meade2007}. These analytic formulae normally assume an infinite half-space geometry and piecewise constant slip. Even so, they are time consuming to derive. To avoid these limitations, the basis for our boundary element formulation is numerical quadrature of the integrals in the full-space elastic Somigliana identity \citep{Cruse1969}. The difficulty in this approach is that straightforward quadrature methods (for example, Gaussian quadrature) assume a smooth integrand. For observation points on or very close to a surface, including the free surface, the integrand is singular. We resolve this problem by evaluating singular integrals as the limit of a sequence of non-singular integrals. This approach is similar to analytical limit-to-the-boundary methods \citep{Sutradhar2008} and series-expansion methods \citep{Klockner2013}. Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived half-space solutions and is free of the singularities present in those solutions. This entirely numerical boundary element method allows for complete flexibility in representing both variable surface topography and complex fault system geometeries. The steep topography of the Longmen Shan rangefront is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.

>>>>

To model both the tectonic and earthquake cycle processes contributing to these interseismic GPS data, we use a quasistatic elastic boundary element software under active development. An important feature of our tool is the inclusion of accurate surface topography in steep mountainous regions or Earth curvature in large-scale problems. Previous boundary element methods in the earth sciences \citep{Thomas1993, Cooke2011, Maerten2014a} have been based on analytic integrations of rectangular or triangular slip surfaces \citep{Okada1992, Meade2007}. These analytic formulae normally assume an infinite half-space geometry and piecewise constant slip. Even so, they are time consuming to derive. To avoid these limitations, the basis for our boundary element formulation is numerical quadrature of the integrals in the full-space elastic Somigliana identity \citep{Cruse1969}. The difficulty in this approach is that straightforward quadrature methods (for example, Gaussian quadrature) assume a smooth integrand. For observation points on or very close to a surface, including the free surface, the integrand is singular. We resolve this problem by evaluating singular integrals as the limit of a sequence of non-singular integrals. This approach is similar to analytical limit-to-the-boundary methods \citep{Sutradhar2008} and series-expansion methods \citep{Klockner2013}. Our tests of relative error show that this quadrature-based method can match, to arbitrary precision, analytically derived half-space solutions and is free of the singularities present in those solutions. This entirely numerical boundary element method allows for complete flexibility in representing both variable surface topography and complex fault system geometeries. The steep topography of the Longmen Shan rangefront is included via the SRTM30\_PLUS digital elevation dataset \citep{Becker2009}.
